Quantifying the overlap of ASVs between taxa

### 2020-10-22 exploratory/2020-10-21-asv-taxa-overlap.Rmd
Randomly subset N genomes from each species, addresses #31

---
title: "Quantifying the overlap of ASVs between taxa"
author: "Pat Schloss"
date: "10/21/2020"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: inline
---

@@

```{r}
# metadata_asv - input data
overlap_data <- metadata_asv %>%
# - focus on taxonomic ranks - kingdom to species, asv, and region
	select(-genome_id, -count, -strain) %>%
# - make data frame tidy
	pivot_longer(cols=c(-region, -asv), names_to="rank", values_to="taxon") %>%
# - remove lines from the data where we don't have a taxonomy
	drop_na(taxon) %>%
# - remove redundant lines 
	distinct() %>%
# for each region and taxonomic rank, group by asvs
	group_by(region, rank, asv) %>%
# - for each asv - count the number of taxa
	summarize(n_taxa = n(), .groups="drop_last") %>%
# - count the number of asvs that appear in more than one taxa
	count(n_taxa) %>%
# - find the ratio of the number of asvs in more than one taxa to the total number of asvs
	summarize(overlap = sum((n_taxa > 1) * n) / sum(n), .groups="drop") %>%
	mutate(rank = factor(rank, levels=c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
	mutate(overlap = 100*overlap)

>>>>

---
title: "Quantifying the overlap of ASVs between taxa"
author: "Pat Schloss"
date: "10/21/2020"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: console
---

@@

```{r}
# set RNG seed to Pat's birthday!
set.seed(19760620)
threshold <- 3

# metadata_asv - species - genome_id
subsample_species <- metadata_asv %>%
	select(genome_id, species) %>%
# return the distinct/unique rows
	distinct() %>%
# group_by the species
	group_by(species) %>%
# slice_sample on each species for N genomes
	slice_sample(n=threshold) %>%
	ungroup()

good_species <- subsample_species %>% 
# count number of genoems in each species
	count(species) %>%
# return species that have N genomes / filter out species with fewer than N genomes
	filter(n == threshold) %>%
	select(species)

# going back to original list of species/genomes and return genome_ids from
# species with N genomes

subsampled_genomes <- inner_join(subsample_species, good_species,  by="species") %>%
	select(genome_id)

# metadata_asv - input data
overlap_data <- inner_join(metadata_asv, subsampled_genomes, by="genome_id") %>%
# - focus on taxonomic ranks - kingdom to species, asv, and region
	select(-genome_id, -count, -strain) %>%
# - make data frame tidy
	pivot_longer(cols=c(-region, -asv), names_to="rank", values_to="taxon") %>%
# - remove lines from the data where we don't have a taxonomy
	drop_na(taxon) %>%
# - remove redundant lines 
	distinct() %>%
# for each region and taxonomic rank, group by asvs
	group_by(region, rank, asv) %>%
# - for each asv - count the number of taxa
	summarize(n_taxa = n(), .groups="drop_last") %>%
# - count the number of asvs that appear in more than one taxa
	count(n_taxa) %>%
# - find the ratio of the number of asvs in more than one taxa to the total number of asvs
	summarize(overlap = sum((n_taxa > 1) * n) / sum(n), .groups="drop") %>%
	mutate(rank = factor(rank, levels=c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
	mutate(overlap = 100*overlap)
